---
title: "SCUC District Subject Assessment Proportions, 1995–2026"
subtitle: "Policycraft longitudinal-data note"
author: "Dan Swart, CPA (ret)"
date: "September 11, 2026"
date-format: long
format:
  html:
    embed-resources: true
    toc: true
    toc-location: right
    code-copy: true
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    df-print: paged
    fig-width: 13
    fig-height: 8
    fig-dpi: 300
editor: source
execute-dir: project
execute:
  echo: false
  message: false
  warning: false
  error: false
  cache: false
---

```{r}
#| label: setup
#| include: false

base::options(scipen = 999)
base::set.seed(123)

locate_output <- function(filename) {
  candidates <- c(
    file.path("outputs", filename),
    file.path("..", "outputs", filename)
  )
  found <- candidates[file.exists(candidates)]
  if (length(found) != 1L) {
    stop("Could not uniquely locate required output: ", filename)
  }
  normalizePath(found, mustWork = TRUE)
}

data_path <- locate_output(
  "scuc_district_subject_assessment_proportions_1995_2026_policycraft.csv"
)
validation_path <- locate_output(
  "scuc_district_subject_assessment_proportions_1995_2026_validation.csv"
)
coverage_path <- locate_output(
  "scuc_district_subject_assessment_proportions_1995_2026_subject_coverage.csv"
)

assessment_data <- utils::read.csv(data_path, check.names = FALSE)
validation <- utils::read.csv(validation_path, check.names = FALSE)
coverage <- utils::read.csv(coverage_path, check.names = FALSE)
```

# Purpose

This analytical dataset contains one row for each academic-year ending year and
subject for Schertz-Cibolo-Universal City ISD. It is formatted for
`Policycraft::launch_longitudinal()` and spans 1995 through 2026. The outcome is
the district-wide proportion of tested students meeting the applicable passing
standard or higher.

Use these fields in Policycraft:

- `date`: date field (integer academic-year ending year);
- `value`: value field (proportion from 0 to 1); and
- `series_label`: grouping field (subject).

Enable percentage display. `subject_sort` provides the intended display order:
Reading, Writing, Mathematics, Science, and Social Studies. The outcome's
success direction is higher is better.

# Coverage

```{r}
#| label: coverage-summary

coverage_display <- coverage[c(
  "subject", "years_with_results", "include_more_than_five_years"
)]
names(coverage_display) <- c(
  "Subject", "Years with results", "Retained"
)
knitr::kable(coverage_display, row.names = FALSE)
```

All five subjects have more than five years with results and are retained. The
CSV uses a complete 1995–2026 year-subject grid so missing administrations are
visible rather than silently dropped.

The modern canonical source contains both 2024–2025 and 2025–2026. Thus, rows
with `date` equal to 2025 and 2026 have been appended for Reading, Mathematics,
Science, and Social Studies. Writing is retained as a historical series but is
not separately reported after its integration into the redesigned Reading
Language Arts assessment.

# Assessment-program labels

Each row identifies the testing program associated with its academic-year
ending year:

| Ending years | `exam_type` | Assessment |
|---|---|---|
| 1995–2002 | TAAS | Texas Assessment of Academic Skills |
| 2003–2011 | TAKS | Texas Assessment of Knowledge and Skills |
| 2012–2026 | STAAR | State of Texas Assessments of Academic Readiness |

The `exam_type` field should be used to identify regime changes, not as evidence
that scores are directly comparable across those changes.

# Calculation

For STAAR years, the dataset uses the requested ratio-of-sums method:

$$
\text{District subject proportion}
=
\frac{\sum_g \text{tests passed}_g}
     {\sum_g \text{tests taken}_g}.
$$

The numerator is the `Satisfactory` count for 2012–2016 and the `Approaches and
Above` count from 2017 onward. Those labels are the applicable passing-level
fields in the canonical source. The numerator and denominator are aggregated
separately across the grades in which each subject was administered before the
ratio is calculated. This avoids averaging grade-level percentages and gives
each tested student the correct weight.

The contributing grades generally are:

- Reading and Mathematics: Grades 3–8;
- Science: Grades 5 and 8;
- Social Studies: Grade 8; and
- Writing: Grades 4 and 7 while separately administered.

For 1995–2011, the available source contains published district-wide subject
percentages but not tests-passed and tests-taken counts. Those values are
therefore retained as reported proportions and identified by
`calculation_method = "reported_district_subject_proportion"`. They were not
recalculated or presented as count-weighted values.

# Important limitations

1. **Assessment regimes are not interchangeable.** TAAS, TAKS, and STAAR used
   different standards, designs, content expectations, and reporting rules.
   A regime boundary can create a discontinuity unrelated to district process
   performance. Analyze within regimes or annotate boundaries prominently.

2. **The population definition changes at 2012.** The STAAR calculation is
   explicitly Grades 3–8. The earlier source describes district-reported tested
   grades and does not include the counts required to reconstruct an identical
   Grades 3–8 population.

3. **The performance label changes within STAAR.** `Satisfactory` is used for
   2012–2016 and `Approaches and Above` thereafter. The source-specific label is
   retained in `performance_standard`.

4. **There is no 2020 result.** Spring 2020 statewide assessments were canceled.
   The dataset retains an explicit missing row for every subject with
   `reporting_status = "not_administered_statewide"`; missing values are not
   zeroes and are not imputed.

5. **Subjects have different tested-grade coverage.** Subject proportions
   combine only the grades in which that subject was administered. Do not
   interpret differences between subjects as if they represent identical sets
   of students or grades.

6. **Writing becomes unavailable as a separate series.** Beginning with the
   2021–2022 academic year, this dataset marks Writing as
   `not_separately_reported`. It should not be filled from the Reading series.

7. **Full precision is retained.** `value` is the analytical proportion;
   `percent` is provided for convenience. Round only for display.

# Validation

```{r}
#| label: validation-results

validation_display <- validation
names(validation_display) <- c("Check", "Passed")
knitr::kable(validation_display, row.names = FALSE)
```

The build asserts a unique `date × subject` key, valid proportion bounds,
reconciliation of every count-derived STAAR value to `tests_passed /
tests_taken`, the 1995–2026 date range, the subject-duration rule, presence of
2025 and 2026 results, and the explicit 2020 gap.

# Provenance and reproducibility

The build script is `R/build_district_subject_assessment_proportions.R`. It
reads these canonical inputs without modifying them:

- `data/SCUC_Snapshots_1995_to_2023_LONG_CANONICAL.csv`; and
- `../SCUC STAAR Longitudinal Data/data/canonical/SCUC_STAAR_Grades_3_8_Annual_2012_2026_LONG_CANONICAL.csv`.

The accompanying lineage CSV records absolute source paths, MD5 checksums,
file modification times, dimensions, and each source's role. The output is a
derived analytical dataset, not a replacement for either canonical source.
